---
title: "bats_analysis"
author: "CarlosVargas"
date: "December 9, 2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(ggplot2)
```

```{r}
myData <- read.table(file="bats.csv", sep=";", skip=3, header=T)
names(myData)
```

## Study of the relationship between brain weight and body mass
```{r}
phyto=myData[(myData$Diet==1),]
```

```{r, echo=FALSE}
ggplot(phyto, aes(x=BOW, y=BRW)) +
  geom_point(alpha=0.5) + geom_line(alpha=0.5) + ggtitle("Total brain weight vs Body mass") +
  scale_y_continuous(trans='log10') + scale_x_continuous(trans='log10')
```

```{r}
reg1 = lm(BRW ~ BOW, data=phyto)
```

```{r}
summary(reg1)
```
### Mathematical form
see notebook

```{r}
anova(reg1)
```


```{r}
plot(reg1$fitted.values, reg1$residuals, xlab="Predicted", ylab="Residuals")
```

```{r}
plot(reg1,4)
which(phyto$BRW>8000)
phytobis=phyto[which(phyto$BRW<8000),]
reg2 = lm(BRW ~ BOW, data=phytobis)
summary(reg2)
```

```{r}
par(mfcol=c(2,2))
plot(reg1)
```


```{r}
par(mfcol=c(2,2))
plot(reg2)
```

By observing diagnostic graphs 2 and 3 for reg1 and reg2, we can see that in the case of graph 2 (Scale-Location), the data points are grouped describe a pattern.

graph 3 comparison: More symetric wrt zero (reg2). reg1 not that symetric.


## 3. Study of the contribution to the total weight of each part of the
brain